```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rhdf5)
library(SingleCellExperiment)
library(Seurat)
library(data.table)

library(dplyr)
library(dbplyr)
library(tidyr)
library(tibble)
library(readr)
library(ggplot2)
library(magrittr)
library(readxl)
library(stringr)
library(testit)
library(knitr)
library(ggpubr)
library(Matrix.utils)
library(S4Vectors)
library(Biobase)
library(immunedeconv)
library(data.tree)
library(rsample)

```

#loading data

```{r}
load("~/Lieselore/R/immune_deconvolution_benchmark/data/schelker/single_cell_schelker.rda")
  valid_cells = pData(single_cell_schelker)$cell_type != "Unknown"
  eset = single_cell_schelker[, valid_cells]
  exprs(eset) = 2^exprs(eset)-1
  pData(eset) = rownames_to_column(pData(eset), "Cell")
  pData(eset) %<>% rename(method_cell_type=cell_type) %>%
    inner_join(immunedeconv::cell_type_map %>% filter(method_dataset == "schelker_single_cell"))
  pbmc_cells = pData(eset)$source != "pbmc"
  eset = eset[,pbmc_cells]  
  assert("no unknown cells left", !any(pData(eset)$cell_type == "Unknown"))

  signature_mel_ova = eset
  
  l <- pData(signature_mel_ova)
  
  counts <- exprs(signature_mel_ova)
  
subset_l <- l%>%
  group_by("cell_type")%>%
  sample_frac(0.70)

cells_subset <- subset_l$Cell

subset_counts <- select(as.data.frame(counts),all_of(cells_subset))
 
rm(single_cell_schelker,counts,l,eset)
```

#splitting dataset in two: 80% training, 20% validation
```{r}
set.seed(999)
split <- initial_split(subset_l,prop = 0.8, strata = cell_type)

training_l <- training(split)
testing_l <- testing(split)

cells_training <- training_l$Cell
cells_testing <- testing_l$Cell

training_counts <- select(as.data.frame(subset_counts), all_of(cells_training))
testing_counts <- select(as.data.frame(subset_counts), all_of(cells_testing))


```


#creating expressionset class -> immunedeconve package
```{r}
training_sce <- SingleCellExperiment(assays = list(counts = training_counts),
                            colData = training_l)

training_p <- colData(training_sce)[,"Cell"]
training_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(training_sce)), groupings = training_p, fun = "sum")))


training_l2 <- training_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

training_pheno <- new("AnnotatedDataFrame", data=training_l2)

training_exp =  ExpressionSet(training_finaldata, training_pheno)


```

#herhalen voor testing

```{r}
testing_sce <- SingleCellExperiment(assays = list(counts = testing_counts),
                            colData = testing_l)

testing_p <- colData(testing_sce)[,"Cell"]
testing_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(testing_sce)), groupings = testing_p, fun = "sum")))


testing_l2 <- testing_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

testing_pheno <- new("AnnotatedDataFrame", data=testing_l2)

testing_exp =  ExpressionSet(testing_finaldata, testing_pheno)

rm(testing_sce)
```

#creating list of all cell types

```{r echo=FALSE}
cell_types = list(
  available=pData(training_exp) %>% select(cell_type) %>% distinct() %>% pull(cell_type),
  cancer=c("Melanoma cell", "Ovarian carcinoma cell"),
  other=c("Cancer associated fibroblast", "Endothelial cell")
)
cell_types$immune_cells = cell_types$available[!cell_types$available %in% c(cell_types$cancer, cell_types$other)]

cell_types$background_cells = c(cell_types$cancer, cell_types$other)
cell_types$n_background_cells = training_exp %>% pData() %>%
  filter(cell_type %in% cell_types$background_cells) %>%
  group_by(cell_type) %>%
  count() %>%
  {'names<-'(.$n, .$cell_type)}
```

## creating pseudobulk samples with TESTING data set
# Average fraction of tumour cells
To obtain representatitive simulated samples, we are interested in the average fraction of tumour cells vs immune cells in a mixture.

```{r tumor-ct-fractions, results='hide', echo=FALSE, fig.cap="proportion of cell types by tumor sample"}
cell_type_table = pData(testing_exp) %>%
  filter(source %in% c("melanoma", "ascites")) %>%
  mutate(sample=paste(source, donor, sep="_")) %>%
  group_by(source, sample, cell_type) %>%
  summarise(n=n()) %>%
  mutate(freq=n/sum(n))

ggplot(cell_type_table, aes(x=sample, y=freq)) +
  geom_bar(aes(fill=cell_type), stat="identity")
  
  
``` 
```{r, echo=FALSE}
cancer_cells = cell_type_table %>%
  filter(cell_type %in% cell_types$cancer)
```

```{r}
cancer_cell_param = MASS::fitdistr(cancer_cells$freq, "normal")
```


The fractions of a sample are randomly assigned in the following procedure:

1. Draw a random tumour cell content from the distribution fitted above
2. The first half of the samples will use melanoma cells, the second half ovarian cancer cells.
3. Assign the remaining fraction (=not cancer cells) randomly to the remaining cell types (B cell, T cell CD8+, Melanoma cell, T cell CD4+ (non-regulatory), Macrophage/Monocyte, T cell regulatory (Tregs), Cancer associated fibroblast, Dendritic cell, Endothelial cell, NK cell, PBMC, Ovarian carcinoma cell)
```{r mixing_make_proportions, echo=FALSE}
set.seed(999)

is_melanoma = c(rep(TRUE, 50), rep(FALSE, 50))

cell_fractions = lapply(is_melanoma, function(melanoma) {
  cancer_fraction = rnorm(1, mean=cancer_cell_param$estimate[1], sd=cancer_cell_param$estimate[2])
  cancer_fraction = ifelse(cancer_fraction < 0, 0, cancer_fraction)
  cancer_fraction = ifelse(cancer_fraction >= .99, .99, cancer_fraction)

  # associate fraction randomly ot Melanoma or Ovarian
  if(melanoma) {
    cancer_fractions = list(cancer_fraction, 0)
  } else {
    cancer_fractions = list(0, cancer_fraction)
  }
  names(cancer_fractions) = cell_types$cancer

  # compute random fractions for other cells
  use_cell_types = c(cell_types$immune_cells, cell_types$other)
  remaining_fraction = 1 - cancer_fraction
  rnd = sample(0:100, length(use_cell_types), TRUE)
  tmp_fractions = (remaining_fraction * rnd) / sum(rnd)
  names(tmp_fractions) = use_cell_types

  c(cancer_fractions, tmp_fractions)
}) %>% bind_rows()

# assert almost equal
assert("all fractions sum up to 1", all(apply(cell_fractions, 1, sum) - rep(1, 100) < 1e-4))
```

```{r}
set.seed(999)
bulk_eset_mel_ova = make_bulk_eset(eset=testing_exp,
                           cell_fractions = cell_fractions,
                           n=500)

fData(bulk_eset_mel_ova) = rownames_to_column(fData(bulk_eset_mel_ova),var = "gene_symbol")
#saveRDS(bulk_eset_mel_ova, "bulk_eset_mel_ova.rda")

bulk_mel_ova = as.data.frame(exprs(bulk_eset_mel_ova))%>%
  rownames_to_column("Gene")



#write.table(bulk_mel_ova, sep = '\t', "bulk_mel_ova.txt", row.names = FALSE, quote = FALSE, col.names = TRUE)

```


#true fractions van simulated samples
```{r}
bulk_eset_mel_ova = read_rds("loading_data/pseudobulk_data/bulk_eset_mel_ova.rda")
fData(bulk_eset_mel_ova) = rownames_to_column(fData(bulk_eset_mel_ova),var = "gene_symbol")
true_mel_ova = pData(bulk_eset_mel_ova)%>%
  t() %>%
  as_tibble(rownames="cell_type")%>%
  gather(sample, true_fraction, -cell_type)



ggplot(true_mel_ova, aes(x=sample, y=true_fraction)) +
  geom_bar(aes(fill=cell_type), stat="identity") +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

# Making single cell reference matrix with TRAINING

```{r}

a <- t(counts(training_sce))

b <- data.frame(colData(training_sce)[,c("Cell","cell_type")],row.names = NULL)%>%
  mutate(Cell_type = as.character(cell_type))%>%
  select(-cell_type)%>%
  column_to_rownames("Cell")

data <- cbind(a,b)

data2 <- rownames_to_column(data, "Cell")%>%
  select(-Cell)%>%
  select(Cell_type, everything())
  

sigmatrix <- as.data.frame(t(data2))
#saveRDS(sigmatrix, "transposed.rds")

finalsigmatrix <- sigmatrix%>%
  unname()

row.names(finalsigmatrix)[1] <- " "

write.table(finalsigmatrix, sep = '\t', "sigmatrix_mel_ova.txt", row.names = TRUE, quote = FALSE)
```


#Cibersortx results

```{r}
devtools::load_all("~/Lieselore/R/Benchmark/immunedeconv")

res_mel_ova = read.delim("~/Lieselore/R/Signature matrix/cibersort_results/CIBERSORTx_mel_ova.txt")%>%
  select(-P.value,-Correlation,-RMSE,-Absolute.score..sig.score.)%>%
  column_to_rownames("Mixture")

cell_types_mel_ova = c("Macrophage/Monocyte" ,"Ovarian carcinoma cell","Cancer associated fibroblast","B cell","T cell regulatory (Tregs)","Dendritic cell",
                       "Endothelial cell","NK cell" ,"Melanoma cell" ,"T cell CD4+ (non-regulatory)","T cell CD8+")
colnames(res_mel_ova) = cell_types_mel_ova

res_mel_ova= res_mel_ova%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "NK cell", "Macrophage/Monocyte", "Dendritic cell","T cell CD4+ (non-regulatory)"
                    ,"T cell regulatory (Tregs)")


res_mel_ova2 = map_result_to_celltypes(res_mel_ova, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method = "Melanoma and ovarian 5.1 signature matrix")
  
                                                                                                                                                                                                                                      
both_mel_ova = inner_join(res_mel_ova2,true_mel_ova,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_mel_ova, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```
#compare to original cibersort from immunedeconve package
```{r}

set_cibersort_binary("~/Lieselore/R/Benchmark/lib/CIBERSORT/CIBERSORT.R")
set_cibersort_mat("~/Lieselore/R/Benchmark/lib/CIBERSORT/LM22.txt")
res_mel_ova_orig =  deconvolute(bulk_eset_mel_ova,"cibersort_abs", tumor = TRUE, scale_mrna = FALSE,
                                column = "gene_symbol") 
                  



  
res_mel_ova_orig2 = map_result_to_celltypes(res_mel_ova_orig, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method = "LM22")

both_mel_ova_orig = inner_join(res_mel_ova_orig2,true_mel_ova,by=c("sample","cell_type"))

ggplot(both_mel_ova_orig, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersort")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```
#creating plot with cibersort and cibersortx results
```{r}
all_results_cibersort_mel_ova= bind_rows(res_mel_ova2,res_mel_ova_orig2)%>%
  arrange(method, cell_type, sample)

results_with_true_mel_ova = inner_join(all_results_cibersort_mel_ova,true_mel_ova,by=c("sample","cell_type"))

results_with_true_mel_ova %>%
  ggplot(aes(x=true_fraction, y=fraction)) +
  facet_grid(method ~ cell_type, scales="free") +
  geom_point() +
  stat_smooth(aes(color=method), method="lm") +
  scale_x_continuous(breaks=c(.2, .3)) +
  stat_cor(aes(label = ..r.label..),size = 3,label.y.npc="top", label.x.npc = "left")+
  scale_color_discrete(name = "Signature matrix")+
  xlab("True fraction")+
  ggtitle("Melanoma and ovarian 5.1")+
  ylab("Fraction")-> p3

ggsave("plot_melova_both_cibersort.png", width = 20, height = 10)
```



#using melanoma signature to deconvolute mel_ova bulk
```{r}
res_mel_ova_melREF = read.delim("~/Lieselore/R/Signature matrix/cibersort_results/CIBERSORTx_melova_melREF.txt")%>%
  select(-P.value,-Correlation,-RMSE,-Absolute.score..sig.score.)%>%
  column_to_rownames("Mixture")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "cancer cell", "Macrophage/Monocyte")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_mel_ova_melREF2 = map_result_to_celltypes(res_mel_ova_melREF, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)

true_mel_ova_wider = pData(bulk_eset_mel_ova)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")


true_mel_ova_melREF =  map_result_to_celltypes(true_mel_ova_wider, cell_types_epic)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,true_fraction,-cell_type)                                                                                                                                                                                                                                   
both_mel_ova_melREF = inner_join(res_mel_ova_melREF2,true_mel_ova_melREF,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_mel_ova_melREF, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```

```{r}
library(patchwork)

p1 / p2 / p3 +
  plot_annotation(tag_levels = 'a')

ggsave("plot_validation cibersortx.png", width = 16, height = 20)
```

