#melanoma 1.1 minor signature matrix

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rhdf5)
library(SingleCellExperiment)
library(Seurat)
library(data.table)

library(dplyr)
library(dbplyr)
library(tidyr)
library(tibble)
library(readr)
library(ggplot2)
library(magrittr)
library(readxl)
library(stringr)
library(testit)
library(knitr)
library(ggpubr)
library(Matrix.utils)
library(S4Vectors)
library(Biobase)
library(immunedeconv)
library(data.tree)
library(rsample)

```

#Melanoma 1.1 
-> simulated samples + signature matrix: splitting into testing and validation


```{r,loading data}

counts <- Read10X_h5("loading_data/single_cell_data/SKCM_GSE72056_expression.h5")

# metadata
l <- read_tsv("loading_data/single_cell_data/SKCM_GSE72056_CellMetainfo_table.tsv", 
  col_types = list(
  `Celltype (malignancy)` = col_factor(c("Immune cells", "Malignant cells", "Stromal cells")),
  `Celltype (major-lineage)` = col_factor(c("B", "CD4Tconv", "CD8Tex","Endothelial", "Fibroblasts","Malignant" ,"Mono/Macro", "Tprolif")),
  `Celltype (minor-lineage)` = col_factor(c("B","CD4Tn", "CD8Tex","Endothelial","Fibroblasts", "Tfh", "Th1","Malignant", "Monocyte", "Tprolif"))))

l <- l%>%
  rename( Malignancy = `Celltype (malignancy)`,
          Major = `Celltype (major-lineage)`, #make_bulk_eset heeft kolom cell_type nodig
          cell_type = `Celltype (minor-lineage)`)


```

splitting dataset in two: 80% training, 20% validation
```{r}
set.seed(999)
split <- initial_split(l,prop = 0.8, strata = cell_type)

training_l <- training(split)
testing_l <- testing(split)

cells_training <- training_l$Cell
cells_testing <- testing_l$Cell

training_counts <- select(as.data.frame(counts), all_of(cells_training))
testing_counts <- select(as.data.frame(counts), all_of(cells_testing))
```

creating expressionset class -> immunedeconve package
```{r}
training_sce <- SingleCellExperiment(assays = list(counts = training_counts),
                            colData = training_l)

training_p <- colData(training_sce)[,"Cell"]
training_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(training_sce)), groupings = training_p, fun = "sum")))


training_l2 <- training_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

training_pheno <- new("AnnotatedDataFrame", data=training_l2)

training_exp =  ExpressionSet(training_finaldata, training_pheno)

```

herhalen voor testing
```{r}
testing_sce <- SingleCellExperiment(assays = list(counts = testing_counts),
                            colData = testing_l)

testing_p <- colData(testing_sce)[,"Cell"]
testing_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(testing_sce)), groupings = testing_p, fun = "sum")))


testing_l2 <- testing_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

testing_pheno <- new("AnnotatedDataFrame", data=testing_l2)

testing_exp =  ExpressionSet(testing_finaldata, testing_pheno)
```

creating list of all cell types
```{r echo=FALSE}
cell_types = list(
  available=pData(training_exp) %>% select(cell_type) %>% distinct() %>% pull(cell_type),
  major=pData(training_exp) %>% select(Major) %>% distinct() %>% pull(Major),
  cancer=c("Malignant"),
  other=c("Endothelial","Fibroblasts")
)
cell_types$immune_cells = as.character(cell_types$available[!cell_types$available %in% c(cell_types$cancer, cell_types$other)])

cell_types$background_cells = c(cell_types$cancer, cell_types$other)
cell_types$n_background_cells = training_exp %>% pData() %>%
  filter(cell_type %in% cell_types$background_cells) %>%
  group_by(cell_type) %>%
  count() %>%
  {'names<-'(.$n, .$cell_type)}
```

creating pseudobulk samples with TESTING data set

Average fraction of tumour cells
```{r tumor-ct-fractions, results='hide', echo=FALSE, fig.cap="proportion of cell types by tumor sample"}
cell_type_table = pData(testing_exp) %>%
  filter(Source %in% "Tumor") %>% #alleen tumor samples selecteren 
  mutate(Sample=paste(Source, Patient, sep="_")) %>%
  group_by(Source, Sample, cell_type) %>%
  summarise(n=n()) %>%
  mutate(freq=n/sum(n))

ggplot(cell_type_table, aes(x=Sample, y=freq)) +
  geom_bar(aes(fill=cell_type), stat="identity")
  
pData(testing_exp) %>%
  group_by(cell_type) %>%
  count()   
```

```{r, echo=FALSE}
cancer_cells = cell_type_table %>%
  filter(cell_type %in% cell_types$cancer)
```

```{r}
cancer_cell_param = MASS::fitdistr(cancer_cells$freq, "normal")
```

```{r mixing_make_proportions, echo=FALSE}

set.seed(999)
is_melanoma = rep(TRUE, 100)

cell_fractions = lapply(is_melanoma, function(melanoma) {
  cancer_fraction = rnorm(1, mean=cancer_cell_param$estimate[1], sd=cancer_cell_param$estimate[2])
  cancer_fraction = ifelse(cancer_fraction < 0, 0, cancer_fraction)
  cancer_fraction = ifelse(cancer_fraction >= .99, .99, cancer_fraction)
  names(cancer_fraction) = cell_types$cancer

  
  # compute random fractions for other cells
  use_cell_types = c(cell_types$immune_cells, cell_types$other)
  remaining_fraction = 1 - cancer_fraction
  rnd = sample(0:100, length(use_cell_types), TRUE)
  tmp_fractions = (remaining_fraction * rnd) / sum(rnd)
  names(tmp_fractions) = use_cell_types

  c(cancer_fraction, tmp_fractions)
}) %>% bind_rows()

# assert almost equal
assert("all fractions sum up to 1", all(apply(cell_fractions, 1, sum) - rep(1, 100) < 1e-4))


```

```{r}
set.seed(999)
bulk_mel_1.1_minor = make_bulk_eset(eset=testing_exp,
                           cell_fractions = cell_fractions,
                           n=500)

pData(bulk_mel_1.1_minor)=pData(bulk_mel_1.1_minor)%>%
  mutate(CD4Tconv = Tfh + CD4Tn + Th1)%>%
  rename(`Mono/Macro` = Monocyte)

fData(bulk_mel_1.1_minor) = rownames_to_column(fData(bulk_mel_1.1_minor),var = "gene_symbol")

saveRDS(bulk_mel_1.1_minor,"loading_data/pseudobulk_data/bulk_melanoma_1.1_minor.rds")



bulk_mel_1.1_minor_txt = as.data.frame(exprs(bulk_mel_1.1_minor))%>%
  rownames_to_column("Gene")



write.table(bulk_mel_1.1_minor_txt, sep = '\t', "loading_data/pseudobulk_data/bulk_mel_1.1_minor.txt", row.names = FALSE, quote = FALSE, col.names = TRUE)

```


Making minor single cell reference matrix with TRAINING
```{r}

a <- t(counts(training_sce))

b <- data.frame(colData(training_sce)[,c("Cell","cell_type")],row.names = NULL)%>%
  mutate(Cell_type = as.character(cell_type))%>%
  select(-cell_type)%>%
  column_to_rownames("Cell")

data <- cbind(a,b)

data2 <- rownames_to_column(data, "Cell")%>%
  select(-Cell)%>%
  select(Cell_type, everything())

data2_nocan <- subset(data2, Cell_type !="Malignant")
  

sigmatrix <- as.data.frame(t(data2_nocan))

finalsigmatrix <- sigmatrix%>%
  unname()

row.names(finalsigmatrix)[1] <- " "

write.table(finalsigmatrix, sep = '\t', "sigmatrix_mel_1.1_minor.txt", row.names = TRUE, quote = FALSE)
```



