 
 
#MELANOMA MAJOR


naamgeving bestanden:

kankertype_bulk_signature
M = Major
m = minor
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rhdf5)
library(SingleCellExperiment)
library(Seurat)
library(data.table)

library(dplyr)
library(dbplyr)
library(tidyr)
library(tibble)
library(readr)
library(ggplot2)
library(magrittr)
library(readxl)
library(stringr)
library(testit)
library(knitr)
library(ggpubr)
library(Matrix.utils)
library(S4Vectors)
library(Biobase)
library(immunedeconv)
library(data.tree)
library(rsample)

```

loading data
```{r}

counts <- Read10X_h5("single_cell_data/SKCM_GSE72056_expression.h5")

## metadata
l <- read_tsv("single_cell_data/SKCM_GSE72056_CellMetainfo_table.tsv", 
  col_types = list(
  `Celltype (malignancy)` = col_factor(c("Immune cells", "Malignant cells", "Stromal cells")),
  `Celltype (major-lineage)` = col_factor(c("B", "CD4Tconv", "CD8Tex","Endothelial", "Fibroblasts","Malignant" ,"Mono/Macro", "Tprolif")),
  `Celltype (minor-lineage)` = col_factor(c("B","CD4Tn", "CD8Tex","Endothelial","Fibroblasts", "Tfh", "Th1","Malignant", "Monocyte", "Tprolif")),
  Sample = col_factor(c("P3_Tumor",  "P3_Normal", "P2_Tumor",  "P2_Normal", "P1_Tumor",  "P1_Normal", "P4_Tumor",  "P4_Normal"))))

l <- l%>%
  rename( Malignancy = `Celltype (malignancy)`,
          cell_type = `Celltype (major-lineage)`, #make_bulk_eset heeft kolom cell_type nodig
          Minor = `Celltype (minor-lineage)`)

```

#splitting dataset in two: 80% training, 20% validation
```{r}
set.seed(999)
split <- initial_split(l,prop = 0.8, strata = cell_type)

training_l <- training(split)
testing_l <- testing(split)

cells_training <- training_l$Cell
cells_testing <- testing_l$Cell

training_counts <- select(as.data.frame(counts), all_of(cells_training))
testing_counts <- select(as.data.frame(counts), all_of(cells_testing))
```

#creating expressionset class -> immunedeconve package
```{r}
training_sce <- SingleCellExperiment(assays = list(counts = training_counts),
                            colData = training_l)

training_p <- colData(training_sce)[,"Cell"]
training_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(training_sce)), groupings = training_p, fun = "sum")))


training_l2 <- training_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

training_pheno <- new("AnnotatedDataFrame", data=training_l2)

training_exp =  ExpressionSet(training_finaldata, training_pheno)

```


#herhalen voor testing

```{r}
testing_sce <- SingleCellExperiment(assays = list(counts = testing_counts),
                            colData = testing_l)

testing_p <- colData(testing_sce)[,"Cell"]
testing_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(testing_sce)), groupings = testing_p, fun = "sum")))


testing_l2 <- testing_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

testing_pheno <- new("AnnotatedDataFrame", data=testing_l2)

testing_exp =  ExpressionSet(testing_finaldata, testing_pheno)
```


#creating list of all cell types
```{r echo=FALSE}
cell_types = list(
  available=pData(training_exp) %>% select(cell_type) %>% distinct() %>% pull(cell_type),
  cancer=c("Malignant"),
  other=c("Endothelial","Fibroblasts")
)
cell_types$immune_cells = as.character(cell_types$available[!cell_types$available %in% c(cell_types$cancer, cell_types$other)])

cell_types$background_cells = c(cell_types$cancer, cell_types$other)
cell_types$n_background_cells = training_exp %>% pData() %>%
  filter(cell_type %in% cell_types$background_cells) %>%
  group_by(cell_type) %>%
  count() %>%
  {'names<-'(.$n, .$cell_type)}
```



## creating pseudobulk samples with TESTING data set
# Average fraction of tumour cells
To obtain representatitive simulated samples, we are interested in the average fraction of tumour cells vs immune cells in a mixture.

```{r tumor-ct-fractions, results='hide', echo=FALSE, fig.cap="proportion of cell types by tumor sample"}
cell_type_table = pData(testing_exp) %>%
  filter(Source %in% "Tumor") %>% #alleen tumor samples selecteren 
  mutate(Sample=paste(Source, Patient, sep="_")) %>%
  group_by(Source, Sample, cell_type) %>%
  summarise(n=n()) %>%
  mutate(freq=n/sum(n))

ggplot(cell_type_table, aes(x=Sample, y=freq)) +
  geom_bar(aes(fill=cell_type), stat="identity")
  
  
```

```{r, echo=FALSE}
cancer_cells = cell_type_table %>%
  filter(cell_type %in% cell_types$cancer)
```

```{r}
cancer_cell_param = MASS::fitdistr(cancer_cells$freq, "normal")
```

The fractions of a sample are randomly assigned in the following procedure:

1. Draw a random tumour cell content from the distribution fitted above
2. The first half of the samples will use melanoma cells, the second half ovarian cancer cells.
3. Assign the remaining fraction (=not cancer cells) randomly to the remaining cell types (B cell, T cell CD8+, Melanoma cell, T cell CD4+ (non-regulatory), Macrophage/Monocyte, T cell regulatory (Tregs), Cancer associated fibroblast, Dendritic cell, Endothelial cell, NK cell, PBMC, Ovarian carcinoma cell)
```{r mixing_make_proportions, echo=FALSE}

set.seed(999)
is_melanoma = rep(TRUE, 100)

cell_fractions = lapply(is_melanoma, function(melanoma) {
  cancer_fraction = rnorm(1, mean=cancer_cell_param$estimate[1], sd=cancer_cell_param$estimate[2])
  cancer_fraction = ifelse(cancer_fraction < 0, 0, cancer_fraction)
  cancer_fraction = ifelse(cancer_fraction >= .99, .99, cancer_fraction)
  names(cancer_fraction) = cell_types$cancer

  
  # compute random fractions for other cells
  use_cell_types = c(cell_types$immune_cells, cell_types$other)
  remaining_fraction = 1 - cancer_fraction
  rnd = sample(0:100, length(use_cell_types), TRUE)
  tmp_fractions = (remaining_fraction * rnd) / sum(rnd)
  names(tmp_fractions) = use_cell_types

  c(cancer_fraction, tmp_fractions)
}) %>% bind_rows()

# assert almost equal
assert("all fractions sum up to 1", all(apply(cell_fractions, 1, sum) - rep(1, 100) < 1e-4))


```



```{r}
set.seed(999)
#bulk_eset_melmaj = make_bulk_eset(eset=testing_exp,
                           cell_fractions = cell_fractions,
                           n=500)

#fData(bulk_eset_melmaj) = rownames_to_column(fData(bulk_eset_melmaj),var = "gene_symbol")
#saveRDS(bulk_eset_melmaj,"bulk_melanoma_major.rds")

bulk_eset_melmaj = read_rds("loading_data/pseudobulk_data/bulk_melanoma_major.rds")

#bulk_mel = as.data.frame(exprs(bulk_eset_melmaj))%>%
  rownames_to_column("Gene")



#write.table(bulk_mel, sep = '\t', "bulk_mel_major.txt", row.names = FALSE, quote = FALSE, col.names = TRUE)

```

#true fractions van simulated samples
```{r}


true_melmaj = pData(bulk_eset_melmaj)%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono/Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t() %>%
  as_tibble(rownames="cell_type")%>%
  gather(sample, true_fraction, -cell_type)



ggplot(true_melmaj, aes(x=sample, y=true_fraction)) +
  geom_bar(aes(fill=cell_type), stat="identity") +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

#testen of epic er oke uit ziet
```{r}
devtools::load_all("~/Lieselore/R/Benchmark/immunedeconv")
res_epic = deconvolute(bulk_eset,"epic", tumor = TRUE, scale_mrna = FALSE, 
                       column = "gene_symbol")

cell_types_epic = c("B cell", "Macrophage/Monocyte",                                                                      "NK cell", "T cell CD4+",
                    "T cell CD8+")
                  

res_epic2 = map_result_to_celltypes(res_epic,cell_types_epic)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)

both_epic = inner_join(res_epic2,true,by=c("sample","cell_type"))

ggplot(both_epic, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```

# Making single cell reference matrix with TRAINING

```{r}

a <- t(counts(training_sce))

b <- data.frame(colData(training_sce)[,c("Cell","cell_type")],row.names = NULL)%>%
  mutate(Cell_type = as.character(cell_type))%>%
  select(-cell_type)%>%
  column_to_rownames("Cell")

data <- cbind(a,b)

data2 <- rownames_to_column(data, "Cell")%>%
  select(-Cell)%>%
  select(Cell_type, everything())
  

sigmatrix <- as.data.frame(t(data2))
#saveRDS(sigmatrix, "transposed.rds")

finalsigmatrix <- sigmatrix%>%
  unname()

row.names(finalsigmatrix)[1] <- " "

write.table(finalsigmatrix, sep = '\t', "sigmatrix_mel_major.txt", row.names = TRUE, quote = FALSE)
```


#Cibersortx results

```{r}
devtools::load_all("~/Lieselore/R/Benchmark/immunedeconv")

res_melanoma_MM = read.delim("~/Lieselore/R/Signature matrix/cibersort_results/CIBERSORTx_melanoma_major_majorREF.txt")%>%
  select(-P.value,-Correlation,-RMSE,-Absolute.score..sig.score.)%>%
  column_to_rownames("Mixture")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "Endothelial cell", "Macrophage/Monocyte","Cancer associated fibroblast")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_melanoma_MM2 = map_result_to_celltypes(res_melanoma_MM, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method = "Melanoma 1.1 signature matrix")
  
                                                                                                                                                                                                                                      
both_mel_MM = inner_join(res_melanoma_MM2,true_melmaj,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_mel_MM, aes(x=true_fraction, y=fraction)) +
  labs(title = "result Cibersortx melanoma + melanoma signature")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()

ggsave("plot_cibersortx_mel_mel.png", width = 20, height = 10)
```


#compare to original cibersort from immunedeconve package
```{r}
set_cibersort_binary("~/Lieselore/R/Benchmark/lib/CIBERSORT/CIBERSORT.R")
set_cibersort_mat("~/Lieselore/R/Benchmark/lib/CIBERSORT/LM22.txt")
res_melanoma_M_orig =  deconvolute(bulk_eset_melmaj,"cibersort_abs", tumor = TRUE, scale_mrna = FALSE, 
                       column = "gene_symbol")


show_cell_types_cibersort = c("B cell", "T cell CD4+", "T cell CD8+", "Macrophage/Monocyte")
  
res_melanoma_M_orig2 = map_result_to_celltypes(res_melanoma_M_orig, show_cell_types_cibersort)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method="LM22")

both_mel_orig = inner_join(res_melanoma_M_orig2,true_melmaj,by=c("sample","cell_type"))

ggplot(both_mel_orig, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersort melanoma + original signature")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()

ggsave("plot_cibersortx_mel_LM22.png", width = 20, height = 10)
```
#creating plot with cibersort and cibersortx results
```{r}
all_results_cibersort= bind_rows(res_melanoma_M_orig2,res_melanoma_MM2)%>%
  arrange(method, cell_type, sample)

results_with_true = inner_join(all_results_cibersort,true_melmaj,by=c("sample","cell_type"))

results_with_true %>%
  ggplot(aes(x=true_fraction, y=fraction)) +
  facet_grid(method ~ cell_type, scales="free") +
  geom_point() +
  stat_smooth(aes(color=method), method="lm") +
  scale_x_continuous(breaks=c(.2, .3)) +
  stat_cor(aes(label = ..r.label..),size = 3,label.y.npc="top", label.x.npc = "left")+
  scale_color_discrete(name = "Signature matrix")+
  xlab("True fraction")+
  ggtitle("Melanoma 1.1")+
  ylab("Fraction") -> p1

ggsave("plot_melanoma_both_cibersort.png", width = 20, height = 10)
```

#bulk lung met melanoma ref
```{r}
res_lung_melREF = read.delim("~/Lieselore/R/Signature matrix/cibersort_results/CIBERSORTx_lung_melREF.txt")%>%
  select(-P.value,-Correlation,-RMSE,-Absolute.score..sig.score.)%>%
  column_to_rownames("Mixture")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "Macrophage/Monocyte")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_lung_melREF2 = map_result_to_celltypes(res_lung_melREF, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method ="Lung")
                                                                                                                                                                                                                                      
both_lung_melREF = inner_join(res_lung_melREF2,true_lung,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_lung_melREF, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```


#using melanoma signature to deconvolute mel_ova bulk
```{r}
res_mel_ova_melREF = read.delim("~/Lieselore/R/Signature matrix/cibersort_results/CIBERSORTx_melova_melREF.txt")%>%
  select(-P.value,-Correlation,-RMSE,-Absolute.score..sig.score.)%>%
  column_to_rownames("Mixture")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_mel_ova_melREF2 = map_result_to_celltypes(res_mel_ova_melREF, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method= "Mel + Ova")

true_mel_ova_wider = pData(bulk_eset_mel_ova)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")


true_mel_ova_melREF =  map_result_to_celltypes(true_mel_ova_wider, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,true_fraction,-cell_type)                                                                                                                                                                                                                                   
both_mel_ova_melREF = inner_join(res_mel_ova_melREF2,true_mel_ova_melREF,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_mel_ova_melREF, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```

#combining melanoma + lung + mel_ova results in 1 plot
```{r}



all_results_melREF =bind_rows(both_lung_melREF,both_mel_ova_melREF,both_mel_MM)%>%
  arrange(method, cell_type, sample)


all_results_melREF %>%
  ggplot(aes(x=true_fraction, y=fraction)) +
  facet_grid(method ~ cell_type, scales="free") +
  geom_point() +
  stat_smooth(aes(color=method), method="lm") +
  scale_x_continuous(breaks=c(.2, .3)) +
  stat_cor()

ggsave("plot_all_melREF.png", width = 20, height = 10)

```


#melanoma signature bekijken

```{r}
melanoma_signature = read.delim("~/Lieselore/R/Signature matrix/signature_matrices/CIBERSORTx_sigmatrix_melanoma.txt")

B = select(melanoma_signature,c(NAME,B))%>%
  arrange(desc(B))

CD8T = select(melanoma_signature,c(NAME,CD8Tex))%>%
  arrange(desc(CD8Tex))

CD4T = select(melanoma_signature,c(NAME,CD4Tconv))%>%
  arrange(desc(CD4Tconv))

Endothelial =  select(melanoma_signature,c(NAME,Endothelial))%>%
  arrange(desc(Endothelial))

Fibroblast = select(melanoma_signature,c(NAME,Fibroblasts))%>%
  arrange(desc(Fibroblasts))

MonoMacro =  select(melanoma_signature,c(NAME,Mono.Macro))%>%
  arrange(desc(Mono.Macro))

NK =  select(melanoma_signature,c(NAME,NK))%>%
  arrange(desc(NK))  


melanoma_sig_genes = c("CD79A","MS4A1","CD79B","CD8B","CD8A","CD4","CD14","CD68","CSF1R","VWF","COL1A2","COL1A1","COL3A1")

write.(melanoma_sig_genes,sep = " ", "melanoma_sig_genes.txt", )
```


#epic
```{r}
devtools::load_all("~/Lieselore/R/Benchmark/immunedeconv")
res_epic = deconvolute(bulk_eset_melmaj,"epic", tumor = TRUE, scale_mrna = FALSE, 
                       column = "gene_symbol")

cell_types_epic = c("B cell", "Macrophage/Monocyte", "T cell CD4+",
                    "T cell CD8+","Endothelial cell","Cancer associated fibroblast")
                  

res_epic2 = map_result_to_celltypes(res_epic,cell_types_epic)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method = "Original signature")

both_epic = inner_join(res_epic2,true_melmaj,by=c("sample","cell_type"))

ggplot(both_epic, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```

#epic met melanoma signature
```{r}

res_melanoma_epic = read.delim("~/Lieselore/R/Signature matrix/epic_results/EPIC_melanoma_melREF.txt")%>%
  select(-otherCells)%>%
  column_to_rownames("sampleID")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "cancer cell", "Macrophage/Monocyte","Endothelial cell", "Cancer associated fibroblast")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_melanoma_epic2 = map_result_to_celltypes(res_melanoma_epic, cell_types_epic)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method= "Melanoma")
  
                                                                                                                                                                                                                                      
both_mel_epic = inner_join(res_melanoma_epic2,true_melmaj,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_mel_epic, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()

```
#creating plot with epic orig en nieuwe epic op melanoma
```{r}
all_results_epic_mel= bind_rows(res_epic2,res_melanoma_epic2)%>%
  arrange(method, cell_type, sample)

results_with_true_epic_mel = inner_join(all_results_epic_mel,true_melmaj,by=c("sample","cell_type"))

results_with_true_epic_mel %>%
  ggplot(aes(x=true_fraction, y=fraction)) +
  facet_grid(method ~ cell_type, scales="free") +
  geom_point() +
  stat_smooth(aes(color=method), method="lm") +
  scale_x_continuous(breaks=c(.2, .3)) +
  stat_cor()

ggsave("plot_melanoma_both_epic.png", width = 20, height = 10)
```



#epic gebruiken met melanoma sig op mel_ova bulk en lung bulk

```{r}

res_lung_melREF_epic = read.delim("~/Lieselore/R/Signature matrix/epic_results/EPIC_lung_melREF.txt")%>%
  select(-otherCells)%>%
  column_to_rownames("sampleID")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "Macrophage/Monocyte","Endothelial cell")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_lung_melREF_epic2 = map_result_to_celltypes(res_lung_melREF_epic,cell_types_epic)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method ="Lung")
  
                                                                                                                                                                                                                                      
both_lung_epic = inner_join(res_lung_melREF_epic2,true_lung,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_lung_epic, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()

```


```{r}

res_melova_melREF_epic = read.delim("~/Lieselore/R/Signature matrix/epic_results/EPIC_mel_ova_melREF.txt")%>%
  select(-otherCells)%>%
  column_to_rownames("sampleID")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "Macrophage/Monocyte","Endothelial cell")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_melova_melREF_epic2 = map_result_to_celltypes(res_melova_melREF_epic, cell_types_epic)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method="Mel + Ova")
  
                                                                                                                                                                                                                                      
both_melova_epic = inner_join(res_melova_melREF_epic2,true_mel_ova_melREF,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_melova_epic, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()

```
#all results epic
```{r}
all_results_epic =bind_rows(both_lung_epic,both_melova_epic,both_mel_epic)%>%
  arrange(method, cell_type, sample)


all_results_epic %>%
  ggplot(aes(x=true_fraction, y=fraction)) +
  facet_grid(method ~ cell_type, scales="free") +
  geom_point() +
  stat_smooth(aes(color=method), method="lm") +
  scale_x_continuous(breaks=c(.2, .3)) +
  stat_cor()

ggsave("plot_all_epic.png", width = 20, height = 10)


```

#EPIC with melanoma sig + new marker genes list
```{r}
res_melanoma_epic_new = read.delim("~/Lieselore/R/Signature matrix/epic_results/EPIC_melanoma_melREF2.txt")%>%
  select(-otherCells)%>%
  column_to_rownames("sampleID")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "cancer cell", "Macrophage/Monocyte","Endothelial cell")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_melanoma_epic_new2 = map_result_to_celltypes(res_melanoma_epic_new, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method= "Melanoma")
  
                                                                                                                                                                                                                                      
both_mel_epic_new = inner_join(res_melanoma_epic_new2,true_melmaj,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_mel_epic_new, aes(x=true_fraction, y=fraction)) +
  labs(title = "")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```
#influence of sig genes

```{r}
#geen B genes:

res_melanoma_epic_noB = read.delim("~/Lieselore/R/Signature matrix/epic_results/EPIC_melanoma_melREF_noB.txt")%>%
  select(-otherCells)%>%
  column_to_rownames("sampleID")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8Tex,
         "Macrophage/Monocyte" = "Mono.Macro",
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant,
         "Cancer associated fibroblast" = Fibroblasts)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")






res_melanoma_epic_noB2 = map_result_to_celltypes(res_melanoma_epic_noB, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method= "Melanoma")
  
                                                                                                                                                                                                                                      
both_mel_epic_noB = inner_join(res_melanoma_epic_noB2,true_melmaj,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both_mel_epic_noB, aes(x=true_fraction, y=fraction)) +
  labs(title = "")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```

