 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rhdf5)
library(SingleCellExperiment)
library(Seurat)
library(data.table)

library(dplyr)
library(dbplyr)
library(tidyr)
library(tibble)
library(readr)
library(ggplot2)
library(magrittr)
library(readxl)
library(stringr)
library(testit)
library(knitr)
library(ggpubr)
library(Matrix.utils)
library(S4Vectors)
library(Biobase)
library(immunedeconv)
library(data.tree)
library(rsample)

```

loading data
```{r}

counts <- Read10X_h5("loading_data/single_cell_data/NSCLC_GSE117570_expression.h5")

## metadata
l <- read_tsv("loading_data/single_cell_data/NSCLC_GSE117570_CellMetainfo_table.tsv", 
  col_types = list(
  `Celltype (malignancy)` = col_factor(c("Immune cells", "Malignant cells", "Stromal cells")),
  `Celltype (major-lineage)` = col_factor(c("B", "CD4Tconv", "CD8T","DC","Endothelial", "Epithelial","Malignant" ,"Mono/Macro", "NK", "Plasma")),
  `Celltype (minor-lineage)` = col_factor(c("B", "CD8Teff","Endothelial","Epithelial", "M1", "M2","Malignant", "Monocyte", "NK", "Plasma", "Th2","pDC")),
  Sample = col_factor(c("P3_Tumor",  "P3_Normal", "P2_Tumor",  "P2_Normal", "P1_Tumor",  "P1_Normal", "P4_Tumor",  "P4_Normal"))))

l <- l%>%
  rename( Malignancy = `Celltype (malignancy)`,
          cell_type = `Celltype (major-lineage)`, #make_bulk_eset heeft kolom cell_type nodig
          Minor = `Celltype (minor-lineage)`)

```

#splitting dataset in two: 80% training, 20% validation
```{r}
set.seed(999)
split <- initial_split(l,prop = 0.8, strata = cell_type)

training_l <- training(split)
testing_l <- testing(split)

cells_training <- training_l$Cell
cells_testing <- testing_l$Cell

training_counts <- select(as.data.frame(counts), all_of(cells_training))
testing_counts <- select(as.data.frame(counts), all_of(cells_testing))
```

#creating expressionset class -> immunedeconve package
```{r}
training_sce <- SingleCellExperiment(assays = list(counts = training_counts),
                            colData = training_l)

training_p <- colData(training_sce)[,"Cell"]
training_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(training_sce)), groupings = training_p, fun = "sum")))


training_l2 <- training_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

training_pheno <- new("AnnotatedDataFrame", data=training_l2)

training_exp =  ExpressionSet(training_finaldata, training_pheno)

```


#herhalen voor testing

```{r}
testing_sce <- SingleCellExperiment(assays = list(counts = testing_counts),
                            colData = testing_l)

testing_p <- colData(testing_sce)[,"Cell"]
testing_finaldata <- as.matrix(t(aggregate.Matrix(t(counts(testing_sce)), groupings = testing_p, fun = "sum")))


testing_l2 <- testing_l%>%
  arrange(Cell)%>%
  column_to_rownames("Cell")

testing_pheno <- new("AnnotatedDataFrame", data=testing_l2)

testing_exp =  ExpressionSet(testing_finaldata, testing_pheno)
```


#creating list of all cell types
```{r echo=FALSE}
cell_types = list(
  available=pData(training_exp) %>% select(cell_type) %>% distinct() %>% pull(cell_type),
  cancer=c("Malignant"),
  other=c("Epithelial", "Endothelial","Plasma")
)
cell_types$immune_cells = as.character(cell_types$available[!cell_types$available %in% c(cell_types$cancer, cell_types$other)])

cell_types$background_cells = c(cell_types$cancer, cell_types$other)
cell_types$n_background_cells = training_exp %>% pData() %>%
  filter(cell_type %in% cell_types$background_cells) %>%
  group_by(cell_type) %>%
  count() %>%
  {'names<-'(.$n, .$cell_type)}
```



## creating pseudobulk samples with TESTING data set
# Average fraction of tumour cells
To obtain representatitive simulated samples, we are interested in the average fraction of tumour cells vs immune cells in a mixture.

```{r tumor-ct-fractions, results='hide', echo=FALSE, fig.cap="proportion of cell types by tumor sample"}
cell_type_table = pData(testing_exp) %>%
  filter(Source %in% "Tumor") %>% #alleen tumor samples selecteren 
  mutate(Sample=paste(Source, Patient, sep="_")) %>%
  group_by(Source, Sample, cell_type) %>%
  summarise(n=n()) %>%
  mutate(freq=n/sum(n))

ggplot(cell_type_table, aes(x=Sample, y=freq)) +
  geom_bar(aes(fill=cell_type), stat="identity")
  
  
```

```{r, echo=FALSE}
cancer_cells = cell_type_table %>%
  filter(cell_type %in% cell_types$cancer)
```

```{r}
cancer_cell_param = MASS::fitdistr(cancer_cells$freq, "normal")
```

The fractions of a sample are randomly assigned in the following procedure:

1. Draw a random tumour cell content from the distribution fitted above
2. The first half of the samples will use melanoma cells, the second half ovarian cancer cells.
3. Assign the remaining fraction (=not cancer cells) randomly to the remaining cell types (B cell, T cell CD8+, Melanoma cell, T cell CD4+ (non-regulatory), Macrophage/Monocyte, T cell regulatory (Tregs), Cancer associated fibroblast, Dendritic cell, Endothelial cell, NK cell, PBMC, Ovarian carcinoma cell)
```{r mixing_make_proportions, echo=FALSE}

set.seed(999)
is_melanoma = rep(TRUE, 100)

cell_fractions = lapply(is_melanoma, function(melanoma) {
  cancer_fraction = rnorm(1, mean=cancer_cell_param$estimate[1], sd=cancer_cell_param$estimate[2])
  cancer_fraction = ifelse(cancer_fraction < 0, 0, cancer_fraction)
  cancer_fraction = ifelse(cancer_fraction >= .99, .99, cancer_fraction)
  names(cancer_fraction) = cell_types$cancer

  
  # compute random fractions for other cells
  use_cell_types = c(cell_types$immune_cells, cell_types$other)
  remaining_fraction = 1 - cancer_fraction
  rnd = sample(0:100, length(use_cell_types), TRUE)
  tmp_fractions = (remaining_fraction * rnd) / sum(rnd)
  names(tmp_fractions) = use_cell_types

  c(cancer_fraction, tmp_fractions)
}) %>% bind_rows()

# assert almost equal
assert("all fractions sum up to 1", all(apply(cell_fractions, 1, sum) - rep(1, 100) < 1e-4))


```



```{r}
set.seed(999)
bulk_eset_lung = make_bulk_eset(eset=testing_exp,
                           cell_fractions = cell_fractions,
                           n=500)
fData(bulk_eset_lung) = rownames_to_column(fData(bulk_eset_lung),var = "gene_symbol")
#saveRDS(bulk_eset_lung,"bulk_eset_lung.rds")

bulk_eset_lung = read_rds("loading_data/pseudobulk_data/bulk_eset_lung.rds")
bulk_lung = as.data.frame(exprs(bulk_eset_lung))%>%
  rownames_to_column("Gene")



write.table(bulk_lung, sep = '\t', "bulk_lung_major.txt", row.names = FALSE, quote = FALSE, col.names = TRUE)

```

#true fractions van simulated samples
```{r}

true_lung = pData(bulk_eset_lung)%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8T,
         "NK cell" = NK,
         "Macrophage/Monocyte" = "Mono/Macro",
         "Dendritic cell" = DC,
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant)%>%
  t() %>%
  as_tibble(rownames="cell_type")%>%
  gather(sample, true_fraction, -cell_type)



ggplot(true_lung, aes(x=sample, y=true_fraction)) +
  geom_bar(aes(fill=cell_type), stat="identity") +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

#testen of epic er oke uit ziet
```{r}
devtools::load_all("~/Lieselore/R/Benchmark/immunedeconv")
res_epic = deconvolute(bulk_eset,"epic", tumor = TRUE, scale_mrna = FALSE, 
                       column = "gene_symbol")

cell_types_epic = c("B cell", "Macrophage/Monocyte",                                                                      "NK cell", "T cell CD4+",
                    "T cell CD8+")
                  

res_epic2 = map_result_to_celltypes(res_epic,cell_types_epic)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)

both_epic = inner_join(res_epic2,true,by=c("sample","cell_type"))

ggplot(both_epic, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```

# Making single cell reference matrix with TRAINING

```{r}

a <- t(counts(training_sce))

b <- data.frame(colData(training_sce)[,c("Cell","cell_type")],row.names = NULL)%>%
  mutate(Cell_type = as.character(cell_type))%>%
  select(-cell_type)%>%
  column_to_rownames("Cell")

data <- cbind(a,b)

data2 <- rownames_to_column(data, "Cell")%>%
  select(-Cell)%>%
  select(Cell_type, everything())
  

sigmatrix <- as.data.frame(t(data2))
#saveRDS(sigmatrix, "transposed.rds")

finalsigmatrix <- sigmatrix%>%
  unname()

row.names(finalsigmatrix)[1] <- " "

write.table(finalsigmatrix, sep = '\t', "sigmatrix_lung_major.txt", row.names = TRUE, quote = FALSE)


```

#Cibersortx results

```{r}
devtools::load_all("~/Lieselore/R/Benchmark/immunedeconv")

res_lung_major = read.delim("~/Lieselore/R/Signature matrix/cibersort_results/CIBERSORTx_lung_new.txt")%>%
  select(-P.value,-Correlation,-RMSE,-Absolute.score..sig.score.)%>%
  column_to_rownames("Mixture")%>%
  rename("B cell" = B,
         "T cell CD4+" = CD4Tconv,
         "T cell CD8+" = CD8T,
         "NK cell" = NK,
         "Macrophage/Monocyte" = Mono.Macro,
         "Dendritic cell" = DC,
         "Endothelial cell" = Endothelial,
         "cancer cell" = Malignant)%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")



show_cell_types = c("B cell", "T cell CD4+", "T cell CD8+", "NK cell", "Macrophage/Monocyte", "Dendritic cell")

#res_lung2 = map_result_to_celltypes(res_lung, show_cell_types, method = "random")

res_lung_major2 = map_result_to_celltypes(res_lung_major, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method="Lung 2.1 signature matrix")
  
                                                                                                                                                                                                                                      
both = inner_join(res_lung_major2,true_lung,by=c("sample","cell_type"))
                                                                                                                                                                                                                                

ggplot(both, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```

#compare to original cibersort from immunedeconve package
```{r}

set_cibersort_binary("~/Lieselore/R/Benchmark/lib/CIBERSORT/CIBERSORT.R")
set_cibersort_mat("~/Lieselore/R/Benchmark/lib/CIBERSORT/LM22.txt")
res_lung_orig =  deconvolute(bulk_eset_lung,"cibersort_abs", tumor = TRUE, scale_mrna = FALSE, 
                       column = "gene_symbol")


show_cell_types = c("B cell", "Dendritic cell", "T cell CD4+", "T cell CD8+", "NK cell", "Macrophage/Monocyte")

  
res_lung_orig2 = map_result_to_celltypes(res_lung_orig, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method = "LM22")

both_orig = inner_join(res_lung_orig2,true_lung,by=c("sample","cell_type"))

ggplot(both_orig, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersort")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()
```
#compare to results from cibersortx LM22
```{r}

cell_types_cibersort = res_lung_orig$cell_type
res_lung_LM22 = read.delim("~/Lieselore/R/Signature matrix/cibersort_results/CIBERSORTx_lung_majorLM22.txt")%>%
  select(-P.value,-Correlation,-RMSE,-Absolute.score..sig.score.)%>%
  column_to_rownames("Mixture")

colnames(res_lung_LM22) = cell_types_cibersort

res_lung_LM22 = res_lung_LM22%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("cell_type")

res_lung_LM22_2 = map_result_to_celltypes(res_lung_LM22, show_cell_types)%>%
  rownames_to_column("cell_type")%>%
  gather(sample,fraction,-cell_type)%>%
  mutate(method="LM22")

both_orig_2 = inner_join(res_lung_LM22_2,true_lung,by=c("sample","cell_type"))

                                                                                                                                                                                                                                

ggplot(both_orig_2, aes(x=true_fraction, y=fraction)) +
  labs(title = "Cibersortx")+
  facet_grid(~cell_type, scales="free") +
  geom_point() +
  stat_smooth(method="lm") +
  stat_cor()

```

#creating plot with cibersort and cibersortx results
```{r}
all_results_cibersort_lung= bind_rows(res_lung_major2,res_lung_orig2)%>%
  arrange(method, cell_type, sample)

results_with_true_lung = inner_join(all_results_cibersort_lung,true_lung,by=c("sample","cell_type"))

results_with_true_lung %>%
  ggplot(aes(x=true_fraction, y=fraction)) +
  facet_grid(method ~ cell_type, scales="free") +
  geom_point() +
  stat_smooth(aes(color=method), method="lm") +
  scale_x_continuous(breaks=c(.2, .3)) +
  stat_cor(aes(label = ..r.label..),size = 3,label.y.npc="top", label.x.npc = "left")+
  scale_color_discrete(name = "Signature matrix")+
  xlab("True fraction")+
  ggtitle("Lung 2.1")+
  ylab("Fraction") -> p2

ggsave("plot_lung_both_cibersort.png", width = 20, height = 10)
```

